Deploying and Sharing QSAR Models using Predictive Modeling Markup Language

### 2014-11-24 paper.tex
The uses of a published QSAR model

\documentclass[12pt,letterpaper]{article}
\usepackage{fullpage, times}
\usepackage{ctable}
\begin{document}
\title{Deploying and Sharing QSAR Models using Predictive Modeling Markup Language}
\author{Villu Ruusmann${}^{\dagger}$ and Rajarshi Guha${}^{\ddagger}$\\
${}^{\ddagger}$National Center for Advancing Translational Sciences\\ 9800 Medical Center Drive  Rockville, MD 20850 \\ \\
${}^{\dagger}$Your address }
\date{}
\maketitle
\begin{abstract}
  A nice abstract
\end{abstract}

@@

>>>>

\documentclass[12pt,letterpaper]{article}
\usepackage{fullpage, times}
\usepackage{hyperref}
\usepackage{ctable}
\begin{document}
\title{Deploying and Sharing QSAR Models using Predictive Modeling Markup Language}
\author{Villu Ruusmann${}^{\dagger}$ and Rajarshi Guha${}^{\ddagger}$\\
${}^{\ddagger}$National Center for Advancing Translational Sciences\\ 9800 Medical Center Drive  Rockville, MD 20850 \\ \\
${}^{\dagger}$Your address }
\date{}
\maketitle
\begin{abstract}
  A nice abstract
\end{abstract}

@@

\subsection{Model reproducibility vs. Model reusability}

The main question: if you see a QSAR model in the literature, do you
want to analyze its mathematical properties, or do you want to apply
it to your dataset in order to see how it performs?

One of the principles of the scientific method is that the research
must be repeatable and reproducible. From the QSAR point of view there
are not much technical obstacles in the way. The biggest concern is
using proprietary algorithms and software.

Model reproducibility is about marking down the procedure by which
the model can be derived given the same input data. Theoretical methods
do not have random error component[?]. Thus, all deviations must be
explained by systematic errors, which could be related to choosing the
wrong algorithm, wrong parametrization etc.

Documenting the whole model training and validation process with the
help of electronic laboratory notebooks. Should include both human and
machine readable instructions of the process. The latter can be given 
in the form of an executable script (ie. when executed with appropriate
input data, produces the model and output data). Plenty of options to
choose from (open source):
\begin{itemize}
  \item knitr \url{http://yihui.name/knitr/}
  \item Sweave \url{https://www.stat.uni-muenchen.de/~leisch/Sweave/}
  \item Sage \url{http://www.sagemath.org/}
  \item IPython \url{http://ipython.org/}
\end{itemize}

The ONS approach is to use Wiki pages, which contain embedded R code
and plots. For example, see
\url{http://onschallenge.wikispaces.com/MeltingPointModel002}.

Most of the time, however, the reader in interested in the \emph{final
ending state} of the modeling workflow. That is, the mathematical 
relationship that has been properly validated and whose applicability 
domain has been sufficiently characterized.

Essentially, the representation of a model training process boils down
to representing the training algorithm. The most complete representation
is in application code such as R, Python, MATLAB etc.
The algorithm "works" by applying a finite list of well-defined 
instructions to the initial state. After applying each instruction the
system reaches another state. When the list of instructions has been
exhausted the system has reached its final ending state.
The representation of states is technically much more challenging than 
the representation of algorithms. A state may consist of different
parts. We are only interested in that part, which relates to the 
mathematical relationship. A state can be preserved by taking a snapshot
of the particular memory area. To make reading and writing more easier,
should be formalized as a data structure.
